---
title: Correlations between programming languages and beliefs about programming
author: |
  Pierce Edmiston
  <pedmiston@wisc.edu>
output:
  beamer_presentation:
    theme: metropolis
---
```{r config, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
               cache=TRUE)
read_chunk("languages.R")
devtools::load_all()
```
```{r setup}
```

# Research questions

- How is our ability to solve problems affected by the tools we use to
  solve them?
- How do people solve the same problems using different programming languages?

This data blitz:

- Explore survey responses given by professional and academic programmers.
- We asked them what languages they knew and their beliefs about programming.

# Who cares about programming languages?

- Hundreds of different programming languages, but all are formally equivalent.
- Computer scientists think a lot about the differences between
  languages.
- Can we measure the *impact* of different programming languages on
  individual computer scientists?
  
# Can we measure differences between programmers who use different languages?

Of course! What should we predict and why does it matter?

> [Lisp] has assisted a number of our most gifted fellow humans in thinking previously 
  impossible thoughts. (Dijkstra, 1972)

H1
: Functional languages are more transformative than other language paradigms.

H2
: Languages with more paradigms are easier to master.

# What languages did they know?

```{r representativeness}
gridExtra::grid.arrange(freq_plot, rank_corr_plot, nrow = 1)
```

# Functional programming languages

```{r functional-languages}
crotchet::draw_svg("imgs/functional-languages.svg")
```

# Programming languages have many paradigms

```{r python-paradigms}
crotchet::draw_svg("imgs/python-paradigms.svg")
```

# Functional versus imperative languages

```{r functional-v-imperative}
crotchet::draw_svg("imgs/functional-v-imperative.svg")
```

# Functional versus object-oriented languages

```{r functional-v-object-oriented}
crotchet::draw_svg("imgs/functional-v-object.svg")
```

# Full Venn Diagram

```{r functional-v-object-v-imperative}
crotchet::draw_svg("imgs/functional-v-object-v-imperative.svg")
```