---
title: Divergence
output: 
  html_document:
    runtime: shiny
---
```{r config, include=FALSE}
library(knitr)
library(shiny)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
               fig.width=4, fig.height=4, dpi=144,
               cache=FALSE)
```
```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(magrittr)

library(programmingquestionnaire)
data("questionnaire")
data("languages")
data("questions")

top20_languages <- languages %>%
  count(language_name) %>%
  top_n(20, n) %>%
  .$language_name

question_names <- c(
  "cr1", "cp1",
  "repo1","rec1", "cfo1", "cfo2",
  "env1", "env2", "inter1", "inter2",
  # agreement w/ other questions
  "oss1", "fp3", "fp4", "psych",
  # multiple choice questions
  "fp1", "pipe1", "computer"
)

most_proficient_languages <- languages %>%
  group_by(subj_id) %>%
  filter(proficiency > 4 | proficiency == max(proficiency)) %>%
  select(subj_id, language_name)

questionnaire_by_language <- questionnaire %>%
  select(subj_id, contains("_int")) %>%
  rename_all(funs(gsub("_int", "", .))) %>%
  select(subj_id, !!!question_names) %>%
  gather(question_name, response, -subj_id) %>%
  drop_na() %>%
  merge(most_proficient_languages) %>%
  as_data_frame() %>%
  filter(language_name %in% top20_languages)
```

**Which multiple choice questions do people who know different programming languages diverge on the most?**

This question is difficult to answer because each person knows multiple programming languages,
and not all at the same level of proficiency. As a result:

1. Responses are only counted toward programming languages with high proficiency (>= 4). If a person
   did not report any programming languages with proficiency >= 4, the most proficient language was used.
2. Only the top 20 most commonly known languages are reported.

**Divergence** was calculated as the difference between the highest and lowest average response
by people who know different languages on a particular question. The results are reported in the
table below. The distributions of responses by each language can be viewed in the ridgelines
plot below the table.

```{r most-diverging}
renderTable({
  questionnaire_by_language %>%
    split(.$question_name) %>%
    map_dfr(function(question_by_language) {
      range_ <- question_by_language %>%
        group_by(language_name) %>%
        summarize(response = mean(response)) %>%
        .$response %>%
        range()
      
      data_frame(question_name = question_by_language$question_name[[1]], divergence = range_[2] - range_[1])
    }) %>%
    arrange(desc(divergence))
  })
```

```{r single-question}
selectInput(
  inputId = "question_name",
  label = "Question",
  choices = question_names,
  selected = 1
)

renderText({
  filter(questions, question_name == input$question_name)$question_text
})

renderPlot({
  question_by_language <- filter(questionnaire_by_language, question_name == input$question_name)
  
  language_means <- question_by_language %>%
    group_by(language_name) %>%
    summarize(response = mean(response))
  
  ordered_languages <- language_means %>%
    arrange(response) %>%
    .$language_name
  
  question_by_language$language_ordered <- factor(question_by_language$language_name, levels = ordered_languages)
  language_means$language_ordered <- factor(language_means$language_name, levels = ordered_languages)
  
  dist_plot <- ggplot(question_by_language) +
    aes(response, language_ordered) +
    geom_density_ridges(bandwidth = 0.5) +
    geom_segment(data = language_means,
                 aes(x = response, xend = response, y = as.numeric(language_ordered), yend = as.numeric(language_ordered) + .9)) +
    ylab("") +
    theme_ridges()
  
  ylim <- max(count(question_by_language, language_ordered)$n)
  
  hist_plot <- ggplot(question_by_language) +
    aes(language_ordered) +
    geom_histogram(stat = "count") +
    stat_count(geom = "text", aes(label = ..count..), hjust = -0.5) +
    labs(x = "") +
    theme_ridges() +
    coord_flip(ylim = c(0, ylim + 5))
  
  dist_plot + hist_plot +
    plot_layout(widths = c(0.6, 0.4))
})
```
